---
title: "Wealth Inequality After Taxes"
description: |
  Analysis of Wealth Inequality data from TidyTuesday
author: Minsong Kim
date: September 24, 2025
format: 
  html:
    code-fold: true
    code-summary: "Code"
---

```{r}
#| include: false
library(purrr)
tuesdata <- tidytuesdayR::tt_load(2025, week = 31)

income_inequality_processed <- tuesdata |>
  pluck("income_inequality_processed")
income_inequality_raw <- tuesdata |>
  pluck("income_inequality_raw")
```

```{r}
#| include: false
# Clean data *and* sample instructions for use in R provided by Our World in
# Data via https://ourworldindata.org/income-inequality-before-and-after-taxes.
# Minimal cleaning applied.

library(tidyverse)
library(jsonlite)

income_inequality_raw <- readr::read_csv(
  "https://ourworldindata.org/grapher/inequality-of-incomes-before-and-after-taxes-and-transfers-scatter.csv?v=1&csvType=full&useColumnShortNames=true"
) |>
  dplyr::mutate(
    Year = as.integer(Year)
  )

income_inequality_processed <- readr::read_csv(
  "https://ourworldindata.org/grapher/gini-coefficient-before-and-after-tax-lis.csv?v=1&csvType=full&useColumnShortNames=true"
) |>
  dplyr::mutate(
    Year = as.integer(Year)
  )

# Download metadata for use in construction of data dictionaries.
metadata_raw <- jsonlite::fromJSON(
  "https://ourworldindata.org/grapher/inequality-of-incomes-before-and-after-taxes-and-transfers-scatter.metadata.json?v=1&csvType=full&useColumnShortNames=true"
)

metadata_processed <- jsonlite::fromJSON(
  "https://ourworldindata.org/grapher/gini-coefficient-before-and-after-tax-lis.metadata.json?v=1&csvType=full&useColumnShortNames=true"
)

```

The visual below shows the change in Gini Index across selected countries after taxes and government benefits over multiple years. The change was calculated by subtracting the post-tax Gini Index from the pre-tax Gini Index. Since a higher value of Gini Index indicates a higher level of wealth inequality, a positive change means a lower level of wealth inequality post-taxes and benefits.

```{r}
#| message: false
#| warning: false
#| fig-alt: Bar charts showing year-to-year changes in the Gini index for Belgium, Iceland, Luxembourg, and the United States, faceted by country from roughly 1970–2023.
#| fig-width: 25
income_inequality_processed |>
  mutate(gini_change = gini_mi_eq - gini_dhi_eq) |>
  filter(Code %in% c("USA", "BEL", "LUX", "ISL")) |>
  group_by(Entity) |>
  ggplot(aes(x = Year, y = gini_change)) +
    geom_bar(stat = "identity") +
    facet_grid(~ Entity, scales = "free_y") +
    labs(
      y = "Gini Index Change",
      x = "Year",
      Title = "Yearly Gini Index Change Across Selected Countries"
    ) +
    theme(
    text = element_text(size = 40),      
    axis.title = element_text(size = 40),
    axis.text = element_text(size = 20)
  ) 

```

Credit:\
Data Processed By:

<https://ourworldindata.org>

Sources:

<https://www.lisdatacenter.org/our-data/lis-database/>

<https://www.oecd.org/en/data/datasets/income-and-wealth-distribution-database.html>

<https://public.yoda.uu.nl/geo/UU01/AEZZIT.html>

<https://www.gapminder.org/data/documentation/gd003/>

<https://population.un.org/wpp/downloads?folder=Standard%20Projections&group=Most%20used>

<https://github.com/open-numbers/ddf--gapminder--systema_globalis>

TidyTuesday Link:

<https://github.com/rfordatascience/tidytuesday/blob/main/data/2025/2025-08-05/readme.md>

Data Curated By:\
Joe Harmon and the Data Science Learning Community

<https://github.com/jonthegeek>
