---
title: "Pokemon"
description: |
  Analysis of Pokemon data from TidyTuesday
author: Minsong Kim
date: December 9, 2025
format: 
  html:
    code-fold: true
    code-summary: "Code"
---

```{r}
#| include: false
tuesdata <- tidytuesdayR::tt_load(2025, week = 13)
pokemon_df <- tuesdata$pokemon_df
library(tidyverse)
library(scales)
```

\
Every pokemon has a list of base stats that correspond to their abilities. A pokemon with a base speed stat of 10 will be faster than a pokemon with a base speed state of 5. Base stat totals are one way to roughly measure a pokemon's total power level, by adding up the stat totals across all their stats. Generally, if a pokemon with a higher base stat total will be stronger than a pokemon with a lower one. A pokemon's egg group also classifies what kind of pokemon it is. A dragon egg group pokemon will be pokemon that is dragon-like in some way, while a flying egg group pokemon might be a bird of some sort. I also created another measurement called size index, which takes a pokemon's height and weight, and multiplies them together to create a unit of measurement that measures both the size and the density of a pokemon roughly. I then plotted every pokemon by their size index versus their base stat totals to see if there was a correlational relationship between the two. I also colored each pokemon by their primary egg group and created a regression line for each egg group so we can see a pattern for each individual egg group and compare them with one another.

```{r}
#| message: false
#| fig-alt: Scatterplot of Pokémon size index versus base stat total, with points colored by egg group. Each egg group has its own regression line, showing generally positive relationships between size and base stats.
#| fig-width: 20
#| fig-height: 10
pokemon_df |>
  mutate(bst = hp + attack + defense + special_attack + special_defense + speed) |>
  mutate(size_index = height * weight)|>
  
  ggplot(aes(x = size_index, y = bst, color = egg_group_1)) +
    geom_point(size = 2.0) +
    scale_x_log10(labels = comma) +
    geom_smooth(method = "lm", linewidth = 1.2, se = FALSE) +
    labs(
      x = "Size Index (Height * Weight)",
      y = "Base Stat Total",
      title = "Size Index vs. Base Stat Total",
      color = "Egg Group"
    ) +
    theme(
    text = element_text(size = 40),      
    axis.title = element_text(size = 40),
    axis.text = element_text(size = 30)
  )
```

We can see that a positive relationship between the size index and base stat total exists for every egg group, but the steepness of the relationship changes depending on the egg group. I think on a design level, that makes sense. You want pokemon who are super big to be stronger than a pokemon the size of your hand generally; otherwise the design contradicts the reality of the strength of the pokemon, as we humans expect big things to be proportionally stronger than small things.

Credit:\
Dataset Original Source:

<https://bulbapedia.bulbagarden.net/wiki/Main_Page>

<https://github.com/phalt/pokeapi>

<https://pokemon-uranium.fandom.com/wiki/Main_Page>

<https://github.com/HybridShivam/Pokemon>

TidyTuesday Page:

<https://github.com/rfordatascience/tidytuesday/tree/main/data/2025/2025-04-01>

Dataset Curated by Frank Hull:\
<https://github.com/frankiethull>
